PloidySelection_Figs

### 2018-12-27 scripts/PloidySelection_Figs.Rmd
Some figure additions

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
library(assertthat)
library(tidyverse)
library(PopGenome)
library(stringr)
library(data.table)
library(gridExtra)

@@

Allele frequency trajectories plot (dips and tets)
```{r}
# Generate allele frequency trajectory based on deterministic recursion equations   
getTraj = function(s, dom, ploidy, start_freq, N = -9, end_freq = 0.99, maxGens = 9999, maxTries = 10){
  dom_coeffs = getDomCoefs(dom, ploidy)
  attempts = 0
  gen = 0
  p = start_freq
  k = ploidy
  while (attempts <= maxTries & p < end_freq & gen <= maxGens){
    df = data.frame("s" = s, "dom" = dom, "gen" = 0, "freq" = start_freq, "dp1" = 0, "w.bar" = 0, "ploidy" = ploidy)
    p = start_freq
    dp1 = 0
    gen = 0
    het_fits = 1 + dom_coeffs * s
    fits = c(1, het_fits, 1 + s) / (1 + s)
    while (p < end_freq & p > 0 & gen <= maxGens){
      q = 1 - p
      i = seq(0, k)
      Num = sum(choose(k, i) * ((k - i) / k) * (p ^ (k - i)) * (q ^ i) * rev(fits))
      Den = sum(choose(k, i) * (p ^ (k - i)) * (q ^ i) * rev(fits))
      p_prime = Num / Den
      if (N != -9){
        p_prime = sum(rbinom(N, k, p_prime)) / (k * N) 
      }
      df = rbind(df, c(s, dom, gen, p, dp1, Den, k))
      dp1 = p_prime - p
      p = p_prime
      gen = gen + 1
    }
    df = rbind(df, c(s, dom, gen, p, dp1, Den, k))
    attempts = attempts + 1
  }
  if (attempts > maxTries & nrow(df) <= maxGens){
    print(paste("Beneficial allele was lost due to drift for", maxTries, "consecutive attempts"))
  }
  if (nrow(df) > maxGens){
    print(paste("Beneficial allele did not fix before the maximum number of generations (", maxGens, ")."))
  }
  return(df[-1,])
}

@@

Allele frequency trajectories for additive allele with arbitrary ploidy
```{r}
s = 0.1
doms = c(0, 0.1, 0.3, 0.5, 0.7, 0.9, 1)
ploidies = c(2, 4, 6, 8)
traj2 = data.frame()
for (i in 1:length(doms)){
  for (j in 1:length(ploidies)){
    traj2 = rbind(traj2, getTraj(s, doms[i], ploidies[j], 0.05))
  }
}

ggplot(traj2, aes(x=gen/100, y=freq, color=as.factor(ploidy))) + geom_line() + scale_color_manual(name="Ploidy",values=c('red','blue','green','purple')) + theme_bw() + xlab("Generation (100's)") + ylab("Allele Frequency") + facet_grid(~dom, scales="free_x")
```

@@

Plot fixation probabilities as a function of ploidy
```{r}
#Fixation probs of diploids as function of N
getFix = function(p, s, N, c){
  num = 1 - exp(-2*c*N*s*p)
  den = 1 - exp(-2*c*N*s)
  return(num / den)
}

p = c(0.0001, 0.001, 0.01, 0.05, 0.1)
N = c(10, 100, 1000, 10000)
s = c(0.001, 0.01, 0.1)
c = c(2, 4, 6, 8)

params = expand.grid(p = p, s = s, N = N, ploidy = c)

params %<>% mutate(fixprob = pmap(list(p, s, N, ploidy), getFix))

params = data.frame("p" = c(0.0001, 0.001, 0.0))
fix = getFix(seq())

```{r}
# Function to loop over all possible combinations of s and h.
simTraj = function(s = c(0.1, 0.01, 0.001), h2 = c(0.25, 0, 1), h1 = c(0.5, 0, 1), h0 = c(0.75, 0, 1), start_freq =  0.05, end_freq = 1.0, maxGen = 9999){
  traj = data.frame()
  for (i in 1:length(s)){
    for (j in 1:length(h1)){
      print(paste("Starting; s =", s[i], ", h =", h1[j]))
      jj = dipTraj(s[i], h1[j], start_freq, end_freq, maxGen)
      kk = tetTraj(s[i], h2[j], h1[j], h0[j], start_freq, end_freq, maxGen)
      traj = rbind(traj, jj)
      traj = rbind(traj, kk)
    }
  }
  return(traj)
}

# Other plots

ggplot(traj, aes(x=gen, y=freq, linetype=as.factor(s), color=ploidy)) + geom_line()+facet_grid(~h1,scales="free_x") + scale_color_discrete()

ggplot(traj[traj$s==0.1,], aes(x=gen, y=var.w, linetype=as.factor(h1), color=ploidy)) + geom_line() + scale_color_manual(name="Ploidy",values=c("red","blue")) + scale_linetype_discrete(name="Dominance", labels=c("Recessive","Additive","Dominant")) + xlab("Generation") + ylab("Variance in Fitness") + theme_bw() + scale_x_log10()

ggplot(traj[traj$s==0.1,], aes(x=gen, y=w.bar, linetype=as.factor(h1), color=ploidy)) + geom_line() + scale_color_manual(name="Ploidy",values=c("red","blue")) + scale_linetype_discrete(name="Dominance", labels=c("Recessive","Additive","Dominant")) + xlab("Generation") + ylab("Mean Fitness") + theme_bw() + scale_x_log10()

ggplot(traj[traj$s==0.1,], aes(x=gen, y=freq, linetype=as.factor(s), color=ploidy)) + geom_line() + scale_color_manual(name="Ploidy",values=c("red","blue")) + scale_linetype_discrete(name="Dominance", labels=c("Recessive","Additive","Dominant")) + ylab("Allele Frequency") + xlab("Generation") + theme_bw()

ggplot(fits[fits$s==0.1,], aes(x=freq, y=var.w, color = as.factor(ploidy), linetype=as.factor(h1))) + geom_line() + scale_color_manual(name="Ploidy",values=c("red","blue")) + scale_linetype_discrete(name="Dominance", labels=c("Recessive","Additive","Dominant")) + xlab("Allele Frequency") + ylab("Variance in Fitness") + theme_bw()

ggplot(fits[fits$s==0.1,], aes(x=freq, y=w.bar, color = as.factor(ploidy), linetype=as.factor(h1))) + geom_line() + scale_color_manual(name="Ploidy",values=c("red","blue")) + scale_linetype_discrete(name="Dominance", labels=c("Recessive","Additive","Dominant")) + xlab("Allele Frequency") + ylab("Mean Fitness") + theme_bw()

@@

# Attempting to implement the diffusion approximations for dips and tets...currently not working
difTraj = function(s, start_freq, end_freq, N, ploidy, max_gens = 10000){
  df = data.frame("s" = s, "gen" = 0, "freq" = start_freq, "ploidy" = ploidy)
  p = start_freq
  gen = 0
  while (p < end_freq & gen < max_gens){
    pq = p * (1 - p)
    mux = 2 * N * s * pq / tanh(2 * N * s * p)
    t = 1 / (ploidy * 2 * N)
    varx = (pq * t) ^ 0.5
    p_prime = p + mux * t
    if (runif(1) >= 0.5){
      p_prime = p_prime + varx
    }
    else{
      p_prime = p_prime - varx
    }
    df = rbind(df, c(s, gen, p_prime, ploidy))
    print(c(p_prime, gen, mux, t, mux * t, varx))
    p = p_prime
    gen = gen + 1
  }
  return(df)
}

#Functions to run mssell
traj_file = "/Users/pmonnahan/Documents/Research/PloidySim/mssel_output/mssel_input/traj_p4_s0.001_N10000_recessive_rep4-4Alt.txt"
traj_file = "/Users/pmonnahan/Documents/Research/PloidySim/testing.txt"

msselRun = function(N, n, traj_file, L = 1000000, mu = 1e-8, r = 1e-8, ploidy = 2, numWindows = 200, slideRate = 0.5, selPos = 0.5, selPop = 2, ms = "/Users/pmonnahan/Documents/Research/code/dmc/mssel_modified/mssel", ms_summarize = "/Users/pmonnahan/Documents/Research/PloidySim/scripts/ms_summarize.R"){ 
  options(scipen=999)
  p = ploidy

  out1 = str_replace(traj_file, ".txt", ".ms.out")
  out2 = str_replace(traj_file, ".txt", ".R.out")

  theta = N * p * mu * L 
  rho = N * p * r * L
  mig_str = "2 0 40 40 0 -ej 0.2825 2 1"

  args1 = paste((2 * p * n), 1, n * p, n * p, traj_file, L * selPos, "-r", rho, L, "-t", theta, "-I", mig_str)

  args2 = paste(out1, out2, L, numWindows, slideRate, Npops, selPop, n * p, n * p)
  print(args1)
  print(args2)
  system2(ms, args1, stdout = out1)

  system2(ms_summarize, args2)

  new_dat = read.table(out2)

  #change names
  return(new_dat)

}

dat = rbind(dat, new_dat)

## Stochastic simulation of selection for a beneficial allele
PloidyForSim = function(ploidy,  N, s, h, start_freq, end_freq, maxGen, maxTries){

  assert_that(ploidy - length(h) == 1)
  attempts = 0

  #Sample individuals to reproduce
  # taken from https://stats.stackexchange.com/questions/67911/how-to-sample-from-a-discrete-distribution
  sampleDist = function(PLOIDY, n, genoFreqs) { 
    sample(x = seq(0, PLOIDY), n, replace = T, prob = genoFreqs)
  }

  #Initalize population
  fits = sort( c(1, 1 + h * s, 1 + s) / (1 + s))
  Pop = rbinom(n = N, size = ploidy, prob = start_freq)
  p = sum(Pop) / (ploidy * N)
  freqs = c(p)

  while(p < end_freq & length(freqs) < maxGen & attempts < maxTries){
    #Retry if stochastic loss of beneficial allele in previous generation
    if ( p == 0 ){
      Pop = rbinom(n = N, size = ploidy, prob = start_freq)
      p = sum(Pop) / (ploidy * N)
      freqs = c(p)
      attempts = attempts + 1
    }

    geno_freq = c( table(factor(Pop, levels = as.character(seq(0, ploidy)))) / N )
    g_prime = ( fits * geno_freq ) / sum(fits * geno_freq) 
    # Make individuals for next generation
    p1 = sampleDist(length(g_prime) - 1, N, g_prime)
    p2 = sampleDist(length(g_prime) - 1, N, g_prime)
    Pop = c()
    for (i in 1:N){
      g1 = sample(c(rep(0, ploidy - p1[i]), rep(1, p1[i])), ploidy / 2, replace = F)
      g2 = sample(c(rep(0, ploidy - p2[i]), rep(1, p2[i])), ploidy / 2, replace = F)
      Pop = c(Pop, sum(g1) + sum(g2))
    }
    p = sum(Pop) / (ploidy * N)
    print(p)
    freqs = c(freqs, p)
  }
  return(freqs)
}

writeTraj = function(file, traj, numPops, selPops, timeScale, trajName = "rep1"){
  gen = seq(1, length(traj) + 1) / timeScale
  traj = rev(traj)
  traj = c(traj, 0)
  anc_traj = rep(0, length(traj))
  fileConn = file(file)
  # writeLines(c(paste("ntraj:", "1"), paste("npop:", numPops), "", paste("n:", length(traj), trajName), paste(gen, traj, anc_traj), sep="\t"), fileConn)
  writeLines(c(paste("ntraj:", "1"), paste("npop:", numPops), paste("n:", length(traj), trajName)), con = fileConn)

  dat = data.frame("gen"=gen, "traj"=traj, "anc"=anc_traj)
  print(dat)
  write.table(dat, file, quote=F, col.names = F, row.names = F, sep="\t", append=T)
  close(fileConn)
}
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
library(assertthat)
library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
library(PopGenome)
library(stringr)
library(data.table)
library(gridExtra)
library(magrittr)
```

@@

Allele frequency trajectories plot (dips and tets)
```{r, echo = FALSE}
# Generate allele frequency trajectory based on deterministic recursion equations   
getTraj = function(s, dom, ploidy, start_freq, N = -9, end_freq = 0.99, maxGens = 9999, maxTries = 10){
  dom_coeffs = getDomCoefs(dom, ploidy)
  attempts = 0
  gen = 0
  p = start_freq
  k = ploidy
  while (attempts <= maxTries & p < end_freq & gen <= maxGens){
    df = data.frame("s" = s, "dom" = dom, "gen" = 0, "freq" = start_freq, "dp1" = 0, "w.bar" = 0, "ploidy" = ploidy)
    p = start_freq
    dp1 = 0
    gen = 0
    het_fits = 1 + dom_coeffs * s
    fits = c(1, het_fits, 1 + s) / (1 + s)
    while (p < end_freq & p > 0 & gen <= maxGens){
      q = 1 - p
      i = seq(0, k)
      Num = sum(choose(k, i) * ((k - i) / k) * (p ^ (k - i)) * (q ^ i) * rev(fits))
      Den = sum(choose(k, i) * (p ^ (k - i)) * (q ^ i) * rev(fits))
      p_prime = Num / Den
      if (N != -9){
        p_prime = sum(rbinom(N, k, p_prime)) / (k * N) 
      }
      df = rbind(df, c(s, dom, gen, p, dp1, Den, k))
      dp1 = p_prime - p
      p = p_prime
      gen = gen + 1
    }
    df = rbind(df, c(s, dom, gen, p, dp1, Den, k))
    attempts = attempts + 1
  }
  if (attempts > maxTries & nrow(df) <= maxGens){
    print(paste("Beneficial allele was lost due to drift for", maxTries, "consecutive attempts"))
  }
  if (nrow(df) > maxGens){
    print(paste("Beneficial allele did not fix before the maximum number of generations (", maxGens, ")."))
  }
  return(df[-1,])
}

@@

Allele frequency trajectories for additive allele with arbitrary ploidy
```{r, echo = FALSE}
s = 0.1
doms = c(0, 0.1, 0.3, 0.5, 0.7, 0.9, 1)
ploidies = c(2, 4, 6, 8)
traj2 = data.frame()
for (i in 1:length(doms)){
  for (j in 1:length(ploidies)){
    traj2 = rbind(traj2, getTraj(s, doms[i], ploidies[j], 0.05))
  }
}

ggplot(traj2, aes(x=gen/1000, y=freq, color=as.factor(ploidy))) + geom_line() + scale_color_manual(name="Ploidy",values=c('red','blue','green','purple')) + theme_bw() + xlab("Generation (1k)") + ylab("Allele Frequency") + facet_wrap(~dom, scales="free_x", nrow = 2)
```

@@

Plot fixation probabilities as a function of ploidy
```{r}
#Fixation probs of diploids as function of N
getFix = function(s, N, c){
  p = 1 / (N * c)
  num = 1 - exp(-2*c*N*s*p)
  den = 1 - exp(-2*c*N*s)
  return(as.numeric(num / den))
}

N = c(10, 20, 50, 100, 500, 1000)
s = c(0.001)
c = c(2, 4, 6, 8)

params = expand.grid(s = s, N = N, ploidy = c)

params %<>% mutate(fixprob = pmap(list(s, N, ploidy), getFix)) %>% unnest()

ggplot(params, aes(y = fixprob, x = N, color = as.factor(ploidy))) + geom_line() + geom_hline(yintercept = 2*s, linetype='dashed') + scale_x_log10() + xlab("Population Size") + ylab("Fixation Probability") + scale_color_discrete(name="Ploidy")
```

OLD
```{r}
# # Function to loop over all possible combinations of s and h.
# simTraj = function(s = c(0.1, 0.01, 0.001), h2 = c(0.25, 0, 1), h1 = c(0.5, 0, 1), h0 = c(0.75, 0, 1), start_freq =  0.05, end_freq = 1.0, maxGen = 9999){
#   traj = data.frame()
#   for (i in 1:length(s)){
#     for (j in 1:length(h1)){
#       print(paste("Starting; s =", s[i], ", h =", h1[j]))
#       jj = dipTraj(s[i], h1[j], start_freq, end_freq, maxGen)
#       kk = tetTraj(s[i], h2[j], h1[j], h0[j], start_freq, end_freq, maxGen)
#       traj = rbind(traj, jj)
#       traj = rbind(traj, kk)
#     }
#   }
#   return(traj)
# }
# 
# # Other plots
# 
# ggplot(traj, aes(x=gen, y=freq, linetype=as.factor(s), color=ploidy)) + geom_line()+facet_grid(~h1,scales="free_x") + scale_color_discrete()
# 
# ggplot(traj[traj$s==0.1,], aes(x=gen, y=var.w, linetype=as.factor(h1), color=ploidy)) + geom_line() + scale_color_manual(name="Ploidy",values=c("red","blue")) + scale_linetype_discrete(name="Dominance", labels=c("Recessive","Additive","Dominant")) + xlab("Generation") + ylab("Variance in Fitness") + theme_bw() + scale_x_log10()
# 
# ggplot(traj[traj$s==0.1,], aes(x=gen, y=w.bar, linetype=as.factor(h1), color=ploidy)) + geom_line() + scale_color_manual(name="Ploidy",values=c("red","blue")) + scale_linetype_discrete(name="Dominance", labels=c("Recessive","Additive","Dominant")) + xlab("Generation") + ylab("Mean Fitness") + theme_bw() + scale_x_log10()
# 
# ggplot(traj[traj$s==0.1,], aes(x=gen, y=freq, linetype=as.factor(s), color=ploidy)) + geom_line() + scale_color_manual(name="Ploidy",values=c("red","blue")) + scale_linetype_discrete(name="Dominance", labels=c("Recessive","Additive","Dominant")) + ylab("Allele Frequency") + xlab("Generation") + theme_bw() 
# 
# ggplot(fits[fits$s==0.1,], aes(x=freq, y=var.w, color = as.factor(ploidy), linetype=as.factor(h1))) + geom_line() + scale_color_manual(name="Ploidy",values=c("red","blue")) + scale_linetype_discrete(name="Dominance", labels=c("Recessive","Additive","Dominant")) + xlab("Allele Frequency") + ylab("Variance in Fitness") + theme_bw()
# 
# ggplot(fits[fits$s==0.1,], aes(x=freq, y=w.bar, color = as.factor(ploidy), linetype=as.factor(h1))) + geom_line() + scale_color_manual(name="Ploidy",values=c("red","blue")) + scale_linetype_discrete(name="Dominance", labels=c("Recessive","Additive","Dominant")) + xlab("Allele Frequency") + ylab("Mean Fitness") + theme_bw()

@@

# Attempting to implement the diffusion approximations for dips and tets...currently not working
# difTraj = function(s, start_freq, end_freq, N, ploidy, max_gens = 10000){
#   df = data.frame("s" = s, "gen" = 0, "freq" = start_freq, "ploidy" = ploidy)
#   p = start_freq
#   gen = 0
#   while (p < end_freq & gen < max_gens){
#     pq = p * (1 - p)
#     mux = 2 * N * s * pq / tanh(2 * N * s * p)
#     t = 1 / (ploidy * 2 * N)
#     varx = (pq * t) ^ 0.5
#     p_prime = p + mux * t
#     if (runif(1) >= 0.5){
#       p_prime = p_prime + varx
#     }
#     else{
#       p_prime = p_prime - varx
#     }
#     df = rbind(df, c(s, gen, p_prime, ploidy))
#     print(c(p_prime, gen, mux, t, mux * t, varx))
#     p = p_prime
#     gen = gen + 1
#   }
#   return(df)
# }
# 
# #Functions to run mssell
# traj_file = "/Users/pmonnahan/Documents/Research/PloidySim/mssel_output/mssel_input/traj_p4_s0.001_N10000_recessive_rep4-4Alt.txt"
# traj_file = "/Users/pmonnahan/Documents/Research/PloidySim/testing.txt"
# 
# msselRun = function(N, n, traj_file, L = 1000000, mu = 1e-8, r = 1e-8, ploidy = 2, numWindows = 200, slideRate = 0.5, selPos = 0.5, selPop = 2, ms = "/Users/pmonnahan/Documents/Research/code/dmc/mssel_modified/mssel", ms_summarize = "/Users/pmonnahan/Documents/Research/PloidySim/scripts/ms_summarize.R"){ 
#   options(scipen=999)
#   p = ploidy
#   
#   out1 = str_replace(traj_file, ".txt", ".ms.out")
#   out2 = str_replace(traj_file, ".txt", ".R.out")
#   
#   theta = N * p * mu * L 
#   rho = N * p * r * L
#   mig_str = "2 0 40 40 0 -ej 0.2825 2 1"
#   
#   args1 = paste((2 * p * n), 1, n * p, n * p, traj_file, L * selPos, "-r", rho, L, "-t", theta, "-I", mig_str)
#   
#   args2 = paste(out1, out2, L, numWindows, slideRate, Npops, selPop, n * p, n * p)
#   print(args1)
#   print(args2)
#   system2(ms, args1, stdout = out1)
# 
#   system2(ms_summarize, args2)
#   
#   new_dat = read.table(out2)
#   
#   #change names
#   return(new_dat)
#   
# }
# 
# dat = rbind(dat, new_dat)
# 
# ## Stochastic simulation of selection for a beneficial allele
# PloidyForSim = function(ploidy,  N, s, h, start_freq, end_freq, maxGen, maxTries){
#   
#   assert_that(ploidy - length(h) == 1)
#   attempts = 0
#   
#   #Sample individuals to reproduce
#   # taken from https://stats.stackexchange.com/questions/67911/how-to-sample-from-a-discrete-distribution
#   sampleDist = function(PLOIDY, n, genoFreqs) { 
#     sample(x = seq(0, PLOIDY), n, replace = T, prob = genoFreqs)
#   }
#   
#   #Initalize population
#   fits = sort( c(1, 1 + h * s, 1 + s) / (1 + s))
#   Pop = rbinom(n = N, size = ploidy, prob = start_freq)
#   p = sum(Pop) / (ploidy * N)
#   freqs = c(p)
#   
#   while(p < end_freq & length(freqs) < maxGen & attempts < maxTries){
#     #Retry if stochastic loss of beneficial allele in previous generation
#     if ( p == 0 ){
#       Pop = rbinom(n = N, size = ploidy, prob = start_freq)
#       p = sum(Pop) / (ploidy * N)
#       freqs = c(p)
#       attempts = attempts + 1
#     }
#     
#     geno_freq = c( table(factor(Pop, levels = as.character(seq(0, ploidy)))) / N )
#     g_prime = ( fits * geno_freq ) / sum(fits * geno_freq) 
#     # Make individuals for next generation
#     p1 = sampleDist(length(g_prime) - 1, N, g_prime)
#     p2 = sampleDist(length(g_prime) - 1, N, g_prime)
#     Pop = c()
#     for (i in 1:N){
#       g1 = sample(c(rep(0, ploidy - p1[i]), rep(1, p1[i])), ploidy / 2, replace = F)
#       g2 = sample(c(rep(0, ploidy - p2[i]), rep(1, p2[i])), ploidy / 2, replace = F)
#       Pop = c(Pop, sum(g1) + sum(g2))
#     }
#     p = sum(Pop) / (ploidy * N)
#     print(p)
#     freqs = c(freqs, p)
#   }
#   return(freqs)
# }
# 
# writeTraj = function(file, traj, numPops, selPops, timeScale, trajName = "rep1"){
#   gen = seq(1, length(traj) + 1) / timeScale
#   traj = rev(traj)
#   traj = c(traj, 0)
#   anc_traj = rep(0, length(traj))
#   fileConn = file(file)
#   # writeLines(c(paste("ntraj:", "1"), paste("npop:", numPops), "", paste("n:", length(traj), trajName), paste(gen, traj, anc_traj), sep="\t"), fileConn)
#   writeLines(c(paste("ntraj:", "1"), paste("npop:", numPops), paste("n:", length(traj), trajName)), con = fileConn)
#   
#   dat = data.frame("gen"=gen, "traj"=traj, "anc"=anc_traj)
#   print(dat)
#   write.table(dat, file, quote=F, col.names = F, row.names = F, sep="\t", append=T)
#   close(fileConn)
# }
